### 2015-05-15 whoop.tex
update
space cutting

\noindent\textbf{Translation for Verification }
%
Next, the driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.

SMACK then translates the driver from LLVM-IR to Boogie IVL, which is the input language of \whoop. An important feature of SMACK is that it leverages the pointer-alias analyses of LLVM to efficiently model the heap manipulation operations of C programs in Boogie IVL. This means that \whoop does not need to directly deal with pointers and alias analysis, a hard problem on its own, allowing us to reuse robust existing techniques and focus instead on verification efforts.

@@

\noindent\textbf{Identifying Locks }
%
When the instrumentation phase begins, \whoop performs an inter-procedural static analysis (on the Boogie IVL source code of each entry point) to identify all available locks and rewrite each one (both at declaration and at all access sites) to a unique constant Boogie variable. The reason behind this transformation is that representing all locks statically, instead of their original SMACK pointer-based representation, helps \whoop perform various internal instrumentation and optimization passes.
%
Currently, \whoop only supports mutexes and spinlocks that are available in the Linux kernel APIs. However, it is relatively easy to enhance our tool with knowledge of other locking primitives. If \whoop cannot infer a lock, e.g. because it was created dynamically or was indexed from an array of locks, it will exit with a warning. The are two reasons behind this: (i) it is arguably hard to detect such locks using static analysis; and (ii) a small, fixed number of locks is advocated by Linux experts as good practice when developing drivers~\cite{corbet2005linux}.

>>>>

\noindent\textbf{Translation for Verification }
%
Next, the driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.
%
SMACK then translates the driver from LLVM-IR to Boogie IVL, which is the input language of \whoop. An important feature of SMACK is that it leverages the pointer-alias analyses of LLVM to efficiently model the heap manipulation operations of C programs in Boogie IVL. This means that \whoop does not need to directly deal with pointers and alias analysis, a hard problem on its own, allowing us to reuse robust existing techniques and focus instead on verification efforts.

@@

\noindent\textbf{Identifying Locks }
%
When the instrumentation phase begins, \whoop performs an inter-procedural static analysis (at Boogie IVL level) to identify all available locks and rewrite each one (both at declaration and at all access sites) to a unique constant Boogie variable. The reason behind this transformation is that representing all locks statically, instead of their original SMACK pointer-based representation, helps \whoop perform various internal instrumentation and optimization passes.
%
Currently, \whoop only supports mutexes and spinlocks that are available in the Linux kernel APIs. However, it is relatively easy to enhance our tool with knowledge of other locking primitives. If \whoop cannot infer a lock, e.g. because it was created dynamically or was indexed from an array of locks, it will exit with a warning. The are two reasons behind this: (i) it is arguably hard to detect such locks using static analysis; and (ii) a small, fixed number of locks is advocated by Linux experts as good practice when developing drivers~\cite{corbet2005linux}.
